---
title: "Econometrics III - Assignment 2"
author: "The Explosive Invertibles"
date: "01 Apr 2022"
output:
  pdf_document:
    includes:
      in_header: "preamble.tex"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(forecast)
library(TSstudio)
library(stringr)
library(zoo)
library(ggplot2)
library(tseries)
library(stats)
library(lmtest)
library(Metrics)
library(dynlm)
library(dLagM)
library(tsoutliers)
```

# Question 1

```{r}
# Load data
data <- read.csv("data_assign_p3.csv")
```



```{r}
set.seed(1)
t <- 200
n <- 100

# Create an empty array to hold our time series
x = numeric(size)
simulation <- list()
# Create the first value
x[1] <- rnorm(n = 1, mean = 0, sd = 1)

for (i in 1:n) {
  for (j in 2:t) {
      # Create random noise for this observation
      w.error = rnorm(n = 1, mean = 0, sd = 1)
      # Compute the current value based on the previous value and the current noise
      x[j] <- x[j - 1] + w.error
  }
  simulation[i] <- x
}
# statt Liste, Matrizen
```
2. Provide plots for two stock market time-series at your choice and report 12-period ACF
and PACF functions for those two time series. What does the sample ACF tell you about
the dynamic properties of these stocks?
```{r}
data$DATE<- as.Date(data$DATE, format = "%d/%m/%Y")
```

```{r}
ggplot(data, aes(x = DATE, y = NETFLIX)) + 
  geom_line() 
```
```{r}
ggplot(data, aes(x = DATE, y = APPLE)) + 
  geom_line() 
```

```{r}
Acf(
  data$NETFLIX,
  lag.max = 12
)
```

```{r}
Pacf(
  data$NETFLIX,
  lag.max = 12
)
```
```{r}
Acf(
  data$APPLE,
  lag.max = 12
)
```

```{r}
Pacf(
  data$APPLE,
  lag.max = 12
)
```

DON'T FORGET TO TALK ABOUT CONFIDENCE INTERVALS!

3. Perform an ADF unit-root test for all the 10 time series using the general-to-specific
approach based on the Schwarz Information Criterion (SIC). Report the values of the
ADF test statistics. Is the unit-root hypothesis rejected for any time-series at the 90%
confidence level? Did you expect to reject the unit-root hypothesis for some time-series at
this confidence level? Justify your answer carefully.

```{r}
netflix <- auto.arima(data$NETFLIX, d = 0, D = 0, max.p = 5, max.q = 5, ic = "aic")
```

